Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-04-10 VegOzone.tex
Reformulate reanalysis part and add that to conclusions.

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g. due to wildfires.\\
We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites. We review available ground level ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}

@@

Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMS reanalysis - CAMSRA). The CAMS reanalysis and CAMS Regional Air Quality ensemble mean are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present). All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF. They differ substantially in the assimilated observational ozone data. The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMS reanalysis also includes ozone profiles from satellite retrievals. Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMS Regional Ensemble. The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}. Without too much foreshadowing, the results prevented us from fetching the whole CAMS reanalysis data.

\begin{table}[t]
  \caption{Ozone related reanalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
  \label{tab:ozone_rea}
\begin{tabular}{llcclcc}
\tophline
Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
&              & spatial & temporal & & \\
\middlehline
MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Subset of the data used in this study. The full data set extends into near real time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
\end{table}

@@

Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}). As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed ground-level ozone well. The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle. Ozone concentrations peak already in March followed by a second peak in July. Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b). Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well. The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September. The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$). Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas. In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{ozone_climatologies_portrait}
  \includegraphics[width=8.3cm]{fig07}
  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: CAMS Regional model reanalysis ensemble mean. As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed ground-level ozone well. The MACC reanalysis is in general too low and displays an erroneous seasonal cycle. The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia. The ensemble mean of the Copernicus CAMS regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
  \label{fig:ozone_climatologies}
\end{figure}

There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses. The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level. Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of the latter. The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone. The latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing. However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data. Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citet{ACP:Andersson2017}.

@@

\conclusions[Discussion and conclusions]
\label{sec:conc}
We have exemplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake. We also assessed implications on potential ozone damage risks arising at the advent of climate change. We found that conditions for ozone formation were more favorable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer. Accordingly, peak ozone concentrations occurred more frequently and at higher levels in 2018. This was in particular caused by the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).

>>>>

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g. due to wildfires.\\
We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites. We review available ground level ozone reanalysis products for the region and find only CAMS Regional Air Quality reanalysis ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}

@@

Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA). From Copernicus member institutions a CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble is available for Europe. The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present). All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF. They differ substantially in the assimilated observational ozone data. The MACC reanalysis assimilates only satellite derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals. Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMSRAQ reanalysis ensemble. The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}. In summary, the results prevented us from using any of the global reanalysis data.

\begin{table}[t]
  \caption{Global/refional ozone reanalysis products used in this study.}
  \label{tab:ozone_rea}
\begin{tabular}{llcclcc}
\tophline
Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
&              & spatial & temporal & & \\
\middlehline
MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
CAMSRAQ & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Subset of the data used in this study. The full data set extends into near real time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
\end{table}

@@

Finally, we shall have a look at the climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}). As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed ground-level ozone well. The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle. Ozone concentrations peak already in March followed by a second peak in July. Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b). Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well. The CAMSRA spring peak lags behind observations by $1\,\unit{month}$ and is $5\,\unit{ppb}$ too low, whereas the minimum occurs in January compared to August/September. The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$). The modeled \chem{[O_3]} at Svanvik are slightly higher than at Pallas which might be due to the height at which surface ozone has been evaluated in the model. In contrast, the ensemble mean of the CAMSRAQ reproduces the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but showss on average a slight low-bias ($(2.8\pm 0.5)\,\unit{ppb}$).

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{ozone_climatologies_portrait}
  \includegraphics[width=8.3cm]{fig07}
  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) MACC reanalysis; (b) CAMSRA; (c) CAMSRAQ reanalysis ensemble mean. As representative for the sites, the nearest neighboring grid point has been chosen. The globale products (a), (b) do not reproduce observed ground-level ozone well. The MACC reanalysis (a) is in general too low and displays an erroneous seasonal cycle. The CAMSRA (b) matches better, but does not reproduce the seasonality in northern Fennoscandia. The ensemble mean of CAMSRAQ reanalysis ensemble reproduces the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
  \label{fig:ozone_climatologies}
\end{figure}

There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses. The model enhancements from MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level. Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of the latter. The higher spatial resolution of the regional air quality models used in the CAMSRAQ ensemble compared to the global reanalyses captures small scale depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone. The latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing. However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data. Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citep{ACP:Andersson2017}.

@@

\conclusions[Discussion and conclusions]
\label{sec:conc}
We have studied the ozone seasonality for northern Fennoscandia and the quality of its representation in available global and regional reanalysis products for Europe. We found that both global reanalysis products of ECWMF fail to reproduce observed ozone paterns. Only the CAMSRAQ ensemble reproduces the observed ozone seasonality well, although with a slight low-bias remaining. This confirms that high spartial resosultion and assimilation of in situ observations of ground-level ozone concentrations are a must to constrain reanalysis related to air quality and risk assement ozone damage to vegetation, especially in high latitudes.

We have exemplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake. We also assessed implications on potential ozone damage risks arising at the advent of climate change. We found that conditions for ozone formation were more favorable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer. Accordingly, peak ozone concentrations occurred more frequently and at higher levels in 2018. This was in particular caused by the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
